Probabilistic Forecasts in Hierarchical~Time~Series

### 2018-04-17 ProbabilisticCoherence.tex
Gaussian Example Complete, although some cleaning up of brackets/braces/parenthesis still needed

Recall that
$$
(\bm{S} ~ \vdots~ \bm{R})^{-1} =
\begin{pmatrix}(\bm{R}'_\bot \bm{S})^{-1}\bm{R}'_\bot \\ \cdots \\ (\bm{S}'_\bot \bm{R})^{-1}\bm{S}'_\bot \end{pmatrix} =
\begin{pmatrix}
\bm{G} \\\bm{H}
\end{pmatrix},
$$
where $\bm{G}=(\bm{R}'_\bot \bm{S})^{-1}\bm{R}'_\bot$ and $\bm{H}=(\bm{S}'_\bot \bm{R})^{-1}\bm{S}'_\bot$. Then $q$ can be simplified further as
\begin{align*}
q& =%\frac{1}{(2\pi)^{\frac{n}{2}}\Big|\bm{\hat{\Sigma}}_{t+h}\Big|^{\frac{1}{2}}\Big|\PQ \Big|}
%\exp \Big\{-\frac{1}{2}
\Big[\bt-\PQ\bm{\hat{\mu}}_{t+h}\Big]'%\\[-0.5cm]
\Big[\PQ\bm{\hat{\Sigma}_{t+h}}\PQ'\Big]^{-1}\Big[\bt-\PQ\bm{\hat{\mu}}_{t+h}\Big] %\Big\},
\\[0.5cm]
 & =%\frac{1}{(2\pi)^{\frac{n}{2}}\Big|\PQ\bm{\hat{\Sigma}_{t+h}}\PQ'\Big|^{\frac{1}{2}}}
%\exp \Big\{-\frac{1}{2} 
\begin{pmatrix}\tilde{\bm{b}} - \bm{G}\bm{\hat{\mu}}\\ \tilde{\bm{a}}- \bm{H}\bm{\hat{\mu}}\end{pmatrix}' %&\\[-0.5cm]
%& \hspace*{7cm}
 \Big[\PQ\bm{\hat{\Sigma}_{t+h}}\PQ'\Big]^{-1}\begin{pmatrix}\tilde{\bm{b}} - \bm{G}\bm{\hat{\mu}}\\ \tilde{\bm{a}}- \bm{H}\bm{\hat{\mu}}\end{pmatrix} %\Big\}.
\end{align*}
Since $\Big[\PQ\bm{\hat{\Sigma}_{t+h}}\PQ'\Big] = \begin{pmatrix}
\bm{P}\bm{\hat{\Sigma}_{t+h}}\bm{P}' & \bm{P}\bm{\hat{\Sigma}_{t+h}}\bm{Q}' \\
\bm{Q}\bm{\hat{\Sigma}_{t+h}}\bm{P}' & \bm{Q}\bm{\hat{\Sigma}_{t+h}}\bm{Q}'
\end{pmatrix}$ we have
\begin{align*}
\bm{f_B}(\cdot)&=\frac{1}{(2\pi)^{\frac{n}{2}}\Big|\begin{pmatrix}
  \bm{P}\bm{\hat{\Sigma}_{t+h}}\bm{P}' & \bm{P}\bm{\hat{\Sigma}_{t+h}}\bm{Q}' \\
  \bm{Q}\bm{\hat{\Sigma}_{t+h}}\bm{P}' & \bm{Q}\bm{\hat{\Sigma}_{t+h}}\bm{Q}'
  \end{pmatrix}\Big|^{\frac{1}{2}}}
\exp \Big\{-\frac{1}{2} \begin{pmatrix}\tilde{\bm{b}}_{t+h} - \bm{P}\bm{\hat{\mu}}_{t+h}\\ \tilde{\bm{t}}_{t+h}- \bm{Q}\bm{\hat{\mu}}_{t+h}\end{pmatrix}'\\
&\hspace*{7cm}\begin{pmatrix}
\bm{P}\bm{\hat{\Sigma}_{t+h}}\bm{P}' & \bm{P}\bm{\hat{\Sigma}_{t+h}}\bm{Q}' \\
\bm{Q}\bm{\hat{\Sigma}_{t+h}}\bm{P}' & \bm{Q}\bm{\hat{\Sigma}_{t+h}}\bm{Q}'
\end{pmatrix}^{-1}
\begin{pmatrix}\tilde{\bm{b}}_{t+h} - \bm{P}\bm{\hat{\mu}}_{t+h}\\ \tilde{\bm{t}}_{t+h}- \bm{Q}\bm{\hat{\mu}}_{t+h}\end{pmatrix} \Big\}.
\end{align*}
This is the joint multivariate Gaussian distribution of $(\tilde{\bm{b}}'_{t+h} ~ \vdots~ \tilde{\bm{t}}'_{t+h})'$. Then from \eqref{4.6} and the properties of the multivariate Gaussian distribution, it follows that
\begin{equation}\label{ex:2.1}
\tilde{\bm{f}}(\tilde{\bm{b}}_{t+h})=\frac{1}{(2\pi)^{\frac{n}{2}}\Big|\bm{P}\bm{\hat{\Sigma}_{t+h}}\bm{P}'\Big|^{\frac{1}{2}}}
\exp \Big\{-\frac{1}{2} (\tilde{\bm{b}}_{t+h} - \bm{P}\bm{\hat{\mu}}_{t+h})' (\bm{P}\bm{\hat{\Sigma}_{t+h}}\bm{P}')^{-1}(\tilde{\bm{b}}_{t+h} - \bm{P}\bm{\hat{\mu}}_{t+h}) \Big\}.
\end{equation}

Equation \eqref{ex:2.1} implies $\tilde{\bm{b}}_{t+h} \sim \mathcal{N}(\bm{P}\bm{\hat{\mu}}_{t+h}, \bm{P}\hat{\bm{\Sigma}}_{t+h}\bm{P}')$, where $\bm{P} = (\bm{R}'_\bot \bm{S})^{-1}\bm{R}'_\bot$. Then from \eqref{4.7} it follows that
\begin{equation}\label{eq:gaussianreconciled}
\tilde{\bm{f}}(\tilde{\bm{y}}_{t+h})=\tilde{\bm{f}}(\bm{S}\tilde{\bm{b}}_{t+h}).
\end{equation}
Therefore, the reconciled Gaussian forecast distribution of the whole hierarchy is\\
$\mathcal{N}(\bm{SP}\bm{\hat{\mu}}_{t+h},~ \bm{SP}\hat{\bm{\Sigma}}_{t+h}\bm{P}'\bm{S}')$.

>>>>

Recall that
$$
(\bm{S} ~ \vdots~ \bm{R})^{-1} =
\begin{pmatrix}(\bm{R}'_\bot \bm{S})^{-1}\bm{R}'_\bot \\ \cdots \\ (\bm{S}'_\bot \bm{R})^{-1}\bm{S}'_\bot \end{pmatrix} =
\begin{pmatrix}
\bm{G} \\\bm{H}
\end{pmatrix},
$$
where $\bm{G}=(\bm{R}'_\bot \bm{S})^{-1}\bm{R}'_\bot$ and $\bm{H}=(\bm{S}'_\bot \bm{R})^{-1}\bm{S}'_\bot$. Then $q$ can be simplified further as
\begin{align*}
q& =%\frac{1}{(2\pi)^{\frac{n}{2}}\Big|\bm{\hat{\Sigma}}_{t+h}\Big|^{\frac{1}{2}}\Big|\PQ \Big|}
%\exp \Big\{-\frac{1}{2}
\Big[\bt-\PQ\bm{\hat{\mu}}_{t+h}\Big]'%\\[-0.5cm]
\Big[\PQ\bm{\hat{\Sigma}_{t+h}}\PQ'\Big]^{-1}\Big[\bt-\PQ\bm{\hat{\mu}}_{t+h}\Big] %\Big\},
\\[0.5cm]
 & =%\frac{1}{(2\pi)^{\frac{n}{2}}\Big|\PQ\bm{\hat{\Sigma}_{t+h}}\PQ'\Big|^{\frac{1}{2}}}
%\exp \Big\{-\frac{1}{2} 
\begin{pmatrix}\tilde{\bm{b}} - \bm{G}\bm{\hat{\mu}}\\ \tilde{\bm{a}}- \bm{H}\bm{\hat{\mu}}\end{pmatrix}' %&\\[-0.5cm]
%& \hspace*{7cm}
 \Big[\PQ\bm{\hat{\Sigma}_{t+h}}\PQ'\Big]^{-1}\begin{pmatrix}\tilde{\bm{b}} - \bm{G}\bm{\hat{\mu}}\\ \tilde{\bm{a}}- \bm{H}\bm{\hat{\mu}}\end{pmatrix} %\Big\}.
\end{align*}
%Since $\Big[\PQ\bm{\hat{\Sigma}_{t+h}}\PQ'\Big] = \begin{pmatrix}
%\bm{P}\bm{\hat{\Sigma}_{t+h}}\bm{P}' & \bm{P}\bm{\hat{\Sigma}_{t+h}}\bm{Q}' 
%\bm{Q}\bm{\hat{\Sigma}_{t+h}}\bm{P}' & \bm{Q}\bm{\hat{\Sigma}_{t+h}}\bm{Q}'
%\end{pmatrix}$ we have

Similar manipulations on determinant of the covariance matrix lead to the following expression for the density

\begin{align*}
f(\tilde{\bm{b}},\tilde{\bm{a}})&
=(2\pi)^{-\frac{n}{2}}\Big|
\begin{pmatrix}
  \bm{G}\bm{\hat{\Sigma}}\bm{G}' & \bm{G}\bm{\hat{\Sigma}}\bm{H}' \\
  \bm{H}\bm{\hat{\Sigma}}\bm{G}' & \bm{H}\bm{\hat{\Sigma}}\bm{H}'
\end{pmatrix}
\Big|^{-\frac{1}{2}}
\exp \Big\{-\frac{1}{2} \begin{pmatrix}\tilde{\bm{b}} - \bm{G}\bm{\hat{\mu}}\\ \tilde{\bm{a}}- \bm{H}\bm{\hat{\mu}}\end{pmatrix}'\\
&\hspace*{7cm}
\begin{pmatrix}
\bm{G}\bm{\hat{\Sigma}}\bm{G}' & \bm{G}\bm{\hat{\Sigma}}\bm{H}' \\
\bm{H}\bm{\hat{\Sigma}}\bm{G}' & \bm{H}\bm{\hat{\Sigma}}\bm{H}'
\end{pmatrix}^{-1}
\begin{pmatrix}\tilde{\bm{b}} - \bm{G}\bm{\hat{\mu}}\\ \tilde{\bm{a}}- \bm{H}\bm{\hat{\mu}}\end{pmatrix} \Big\}.
\end{align*}
Marginalising out $\tilde{\bm{a}}$, it follows leads to the following bottom level reconciled forecasts. 
\begin{equation}\label{ex:2.1}
f(\tilde{\bm{b}})=(2\pi)^{-\frac{m}{2}}\Big|\bm{G}\bm{\hat{\Sigma}}\bm{G}'\Big|^{\frac{1}{2}}
\exp \Big\{-\frac{1}{2} (\tilde{\bm{b}} - \bm{G}\bm{\hat{\mu}})' (\bm{G}\bm{\hat{\Sigma}}\bm{G}')^{-1}(\tilde{\bm{b}} - \bm{G}\bm{\hat{\mu}}) \Big\}.
\end{equation}

%Equation \eqref{ex:2.1} implies $\tilde{\bm{b}}_{t+h} \sim \mathcal{N}(\bm{P}\bm{\hat{\mu}}_{t+h}, \bm{P}\hat{\bm{\Sigma}}_{t+h}\bm{P}')$, where $\bm{P} = (\bm{R}'_\bot \bm{S})^{-1}\bm{R}'_\bot$. Then from \eqref{4.7} it follows that
%\begin{equation}\label{eq:gaussianreconciled}
%\tilde{\bm{f}}(\tilde{\bm{y}}_{t+h})=\tilde{\bm{f}}(\bm{S}\tilde{\bm{b}}_{t+h}).
%\end{equation}
Which implies that the reconciled probabilistic forecast for the bottom level series is $\tilde{\bm{b}}_{t+h} \sim \mathcal{N}(\bm{P}\bm{\hat{\mu}}_{t+h}, \bm{P}\hat{\bm{\Sigma}}_{t+h}\bm{P}')$.  The reconciled probabilistic forecasts for the whole hierarchy follow a degenerate Gaussian distribution with mean  $\bm{SG}\bm{\hat{\mu}}$ and rank deficient covariance matrix $\bm{SP}\hat{\bm{\Sigma}}_{t+h}\bm{P}'\bm{S}'$.
